
Title: ARG abundance time series data pre-processing
Author: Josefine Ostenfeld Nielsen, National Food Institute, Technical University of Denmark (josefineon@gmail.com)
Date: January 2024
Description: Linear interpolation of data gaps and adjustment of systematic error

________________________________________________________________________________


Installing packages and loading libraries

```{r, message=F, warning=F}
# Sourcing packages and libraries from the setup.R script 
source(file.path("setup.R"))
```


Importing ARG abundance time series data

```{r}
# Reading ARG abundance file
ARG_abundance <- read.table(file.path("data/ARG_abundance.txt"), header = TRUE, sep = "\t")

# Formatting Date column 
ARG_abundance$Date <- as.Date(ARG_abundance$Date)
```

________________________________________________________________________________


Linear interpolation of data gaps in ARG abundance time series 

```{r}
# Defining AMR-classes
amr_class_names <- c(
  "Aminoglycoside",
  "Amphenicol",
  "Beta.lactam",
  "Lincosamide",
  "Macrolide",
  "Pleuromutilin",
  "Tetracycline"
)

# Creating an empty data frame to store interpolated data
ARG_abundance_interpolated <- data.frame(
  Date = as.Date(character(0)),
  amr = numeric(0),
  amr_class = character(0)
)

# Looping through AMR-classes in ARG abundance dataframe and interpolating ARG abundances in data gaps
for (class in amr_class_names) {

  # Filtering each AMR-class
  data <- ARG_abundance %>%
    filter(amr_class == class) %>%
    select(-amr_class)

  # Sequence of dates spanning full time series range
  all_dates <- data.frame(Date = seq(from = as.Date("2015-03-01"), to = as.Date("2022-12-01"), by = "1 month"))

  # Adding ARG abundances to the dates and NA values to dates with missing ARG abundance values
  data <- left_join(all_dates, data, by = "Date")

  # Converting to zoo object
  data_zoo <- zoo(data$amr, order.by = data$Date)

  # Performing linear interpolation to impute missing values
  data_zoo <- na.approx(data_zoo)

  # Creating a data frame with interpolated data for the current AMR class
  data <- data.frame(Date = time(data_zoo), amr_class = class, amr = coredata(data_zoo))

  # Bind the interpolated data for the current AMR class to data_interpolated
  ARG_abundance_interpolated <- bind_rows(ARG_abundance_interpolated, data)
}
```


Adjusting ARG abundance data from 2015-2018 by multiplying with average ratio between months in 2018 and 2019

```{r}
# Data frame to store ratios
ratios_df <- data.frame(
  amr_class = character(),
  Rc = numeric(),
  stringsAsFactors = FALSE
)

# Data frame to store adjusted ARG abundance data
ARG_abundance_adjusted <- data.frame(
  Date = as.Date(character()),
  amr = numeric(),
  amr_class = character(),
  stringsAsFactors = FALSE
)

# Loop to adjust ARG abundance data from 2015-2018 individually for each AMR-class
for (class in amr_class_names) {

  # Filtering AMR-class, 2018 and 2019 data
  amr <- ARG_abundance_interpolated %>%
    filter(amr_class == class) %>%
    select(-amr_class) %>%
    filter(year(Date) > 2017) %>%
    filter(year(Date) < 2020)

  # Selecting FPKM values from 2018 and 2019
  amr_18 <- amr[(1:12), ]
  amr_19 <- amr[(13:24), ]

  # Estimating ratio between FPKM for each month in 2019 and all month in 2018
  amr_19$amr_diff_dec18 <- amr_19$amr/amr_18$amr[12]
  amr_19$amr_diff_nov18 <- amr_19$amr/amr_18$amr[11]
  amr_19$amr_diff_oct18 <- amr_19$amr/amr_18$amr[10]
  amr_19$amr_diff_sep18 <- amr_19$amr/amr_18$amr[9]
  amr_19$amr_diff_aug18 <- amr_19$amr/amr_18$amr[8]
  amr_19$amr_diff_jul18 <- amr_19$amr/amr_18$amr[7]
  amr_19$amr_diff_jun18 <- amr_19$amr/amr_18$amr[6]
  amr_19$amr_diff_maj18 <- amr_19$amr/amr_18$amr[5]
  amr_19$amr_diff_apr18 <- amr_19$amr/amr_18$amr[4]
  amr_19$amr_diff_mar18 <- amr_19$amr/amr_18$amr[3]
  amr_19$amr_diff_feb18 <- amr_19$amr/amr_18$amr[2]
  amr_19$amr_diff_jan18 <- amr_19$amr/amr_18$amr[1]

  # Estimating the mean ratio between the FPKM values each month in 2019 and 2018
  Rc <- mean(apply(amr_19[, 3:14], MARGIN = 2, function(x) mean(x, na.rm = TRUE)))
  
  # Storing results in data frame for each amr class
  ratios_df <- rbind(ratios_df, data.frame(
    amr_class = class,
    Rc = Rc
  ))
  
  # Adjusting ARG abundance 
  ARG_adj <- ARG_abundance_interpolated %>%
    filter(amr_class == class) %>%
    mutate(across(where(is.numeric), ~ if_else(year(Date) < 2019, . * Rc, .)))

  # Storing adjusted data for each AMR-class
  ARG_abundance_adjusted <- bind_rows(ARG_abundance_adjusted, ARG_adj)
}
```



