
Title: Lifetime AMU and ARG abundance time series decomposition and stationarity
Author: Josefine Ostenfeld Nielsen, National Food Institute, Technical University of Denmark (josefineon@gmail.com)
Date: January 2024
Description: Isolating trend component and differencing, stationarity tests and summary statistics

________________________________________________________________________________


Installing packages and loading libraries

```{r, message=F, warning=F}
# Sourcing packages and libraries from the setup.R script
source(file.path("setup.R"))
```


Importing ARG abundance and lifetime AMU data

```{r}
# Reading ARG abundance and lifetime AMU files
ARG_abundance_adjusted <- read.table(file.path("data/ARG_abundance_adj.txt"), header = TRUE, sep = "\t")
lifetime_AMU <- read.table(file.path("data/lifetime_AMU.txt"), header = TRUE, sep = "\t")

# Formatting Date column 
ARG_abundance_adjusted$Date <- as.Date(ARG_abundance_adjusted$Date)
lifetime_AMU$Date <- as.Date(lifetime_AMU$Date)
```

________________________________________________________________________________


Decomposing ARG abundance time series, extracting trend-cycle component, and differencing 

```{r}
# Defining AMR-classes
amr_class_names <- c(
  "Aminoglycoside",
  "Amphenicol",
  "Beta.lactam",
  "Lincosamide",
  "Macrolide",
  "Pleuromutilin",
  "Tetracycline"
)

# Data frame to store ARG abundance trend component 
ARG_abundance_trend <- data.frame(
  Date = as.Date(character()),
  amr_class = character(),
  amr = numeric(),
  stringsAsFactors = FALSE
)

# Data frame to store differenced ARG abundance trend component 
ARG_abundance_trend_diff <- data.frame(
  Date = as.Date(character()),
  amr_class = character(),
  amr = numeric(),
  stringsAsFactors = FALSE
)

# Loop to decompose the ARG abundance time series individually for each AMR-class and extract trend component 
for (class in amr_class_names) {
  
  # Filtering AMR-class
  amr <- ARG_abundance_adjusted %>%
    filter(amr_class == class) %>%
    select(-amr_class)
  
  # Converting into tsibble object (time series Date format)
  amr_ts <- amr %>%
    mutate(Date = yearmonth(Date)) %>%
    as_tsibble(index = Date)
  
  # Decomposing ARG abundance time series (seasonally adjusting and smoothing)
  amr_decomposed <- stl(amr_ts, s.window = "periodic", t.window = 7)
  amr_decomposed <- data.frame(amr_decomposed$time.series)
  
  # Isolating ARG abundance trend
  amr_trend <- bind_cols(Date = amr$Date, amr = amr_decomposed$trend)
  
  # Removing first and last two values with incomplete smoothing 
  amr_trend <- amr_trend %>%
    slice(3:(n()-2)) %>%
    mutate(amr_class = class)
  
  # Saving ARG abundance trend in data frame 
  ARG_abundance_trend <- rbind(ARG_abundance_trend, amr_trend)
  
  # Differencing ARG abundance trend and removing first value (where differencing cannot be computed)
  amr_diff <- data.frame(amr = diff(amr_trend$amr, differences = 1))
  amr_diff <- bind_cols(Date = amr_trend[-(1), 1], amr_diff)
  amr_diff <- amr_diff %>% mutate(amr_class = class)
  
  # Saving differenced ARG abundance trend in data frame 
  ARG_abundance_trend_diff <- rbind(ARG_abundance_trend_diff, amr_diff)
}
```


Decomposing lifetime AMU time series, extracting trend-cycle component, and differencing 

```{r}
# Defining AM-classes
am_class_names <- c(
  "Aminoglycosides",
  "Amphenicols",
  "SimPenicillins",
  "ExtPenicillins",
  "Lincosamides",
  "Macrolides",
  "Pleuromutilins",
  "Tetracyclines"
)

# Data frame to store lifetime AMU trend component 
lifetime_AMU_trend <- data.frame(
  Date = as.Date(character()),
  amr_class = character(),
  amr = numeric(),
  stringsAsFactors = FALSE
)

# Data frame to store differenced lifetime AMU trend component 
lifetime_AMU_trend_diff <- data.frame(
  Date = as.Date(character()),
  amr_class = character(),
  amr = numeric(),
  stringsAsFactors = FALSE
)

# Loop to decompose the lifetime AMU time series individually for each AM-class and extract trend component
for (class in am_class_names) {
  
  # Filtering AM-class
  amu <- lifetime_AMU %>%
    filter(am_group == class) %>%
    select(-am_group)
  
  # Converting into tsibble object (time series Date format)
  amu_ts <- amu %>%
    mutate(Date = yearmonth(Date)) %>%
    as_tsibble(index = Date)
  
  # Decomposing lifetime AMU time series (seasonally adjusting)
  amu_decomposed <- stl(amu_ts, s.window = "periodic")
  amu_decomposed <- data.frame(amu_decomposed$time.series)
  amu_seasadj <- data.frame(amu$amu - amu_decomposed$seasonal)
  
  # Isolating lifetime AMU trend (smoothing)
  amu_trend <- data.frame(amu = SMA(amu_seasadj, n = 3))
  
  # Removing first two values with incomplete smoothing 
  amu_trend <- bind_cols(Date = amu[-(1:2),1], amu = amu_trend[-(1:2),])
  
  # Differencing lifetime AMU trend and removing first value (where differencing cannot be computed)
  amu_diff <- data.frame(amu = diff(amu_trend$amu, differences = 1))
  amu_diff <- bind_cols(Date = amu_trend[-(1), 1], amu_diff)
  
  # Adding AM-class to data
  amu_trend <- amu_trend %>% mutate(am_group = class) 
  amu_diff <- amu_diff %>% mutate(am_group = class) 
  
  # Saving lifetime AMU trend in data frame 
  lifetime_AMU_trend <- rbind(lifetime_AMU_trend, amu_trend)
  
  # Saving differenced lifetime AMU trend in data frame 
  lifetime_AMU_trend_diff <- rbind(lifetime_AMU_trend_diff, amu_diff)
}  
```

________________________________________________________________________________


Testing stationarity of differenced ARG abundance trend time series 

```{r, warning=F, message=F}
# Defining AMR-classes
am_group_names <- c(
  "Aminoglycoside",
  "Amphenicol",
  "Beta.lactam",
  "Lincosamide",
  "Macrolide",
  "Pleuromutilin",
  "Tetracycline"
)

# Data frame to store stationarity test results 
stationarity_results <- data.frame(
  AMGroup = character(),
  Original_KPSS_stat = numeric(),
  Original_KPSS_p_value = numeric(),
  Original_ADF_stat = numeric(),
  Original_ADF_p_value = numeric(),
  Diff1_KPSS_stat = numeric(),
  Diff1_KPSS_p_value = numeric(),
  Diff1_ADF_stat = numeric(),
  Diff1_ADF_p_value = numeric(),
  stringsAsFactors = FALSE
)

# Loop to test stationarity individually for each AMR-class for the ARG abundance time series trend before and after differencing 
for (am in am_group_names) {

  # Filtering AMR-class of ARG abundance trend time series 
  data <- ARG_abundance_trend %>%
    filter(amr_class == am) %>%
    select(-amr_class) 
  
  # Filtering AMR-class of differenced ARG abundance trend time series 
  data_diff <- ARG_abundance_trend_diff %>%
    filter(amr_class == am) %>%
    select(-amr_class) 
  
  # Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test for trend stationarity
  kpss_test <- kpss.test(data$amr)
  kpss_test_diff <- kpss.test(data_diff$amr)
  
  # Augmented-Dickey–Fuller (ADF) test for trend stationarity
  adf_test <- adf.test(data$amr)
  adf_test_diff <- adf.test(data_diff$amr)
  
  # Storing stationarity results in data frame 
  stationarity_results <- rbind(stationarity_results, data.frame(
    AMGroup = am,
    Original_KPSS_stat = round(kpss_test$statistic,2),
    Original_KPSS_p_value = round(kpss_test$p.value,2),
    Original_ADF_stat = round(adf_test$statistic,2),
    Original_ADF_p_value = round(adf_test$p.value,2),
    Diff1_KPSS_stat = round(kpss_test_diff$statistic,2),
    Diff1_KPSS_p_value = round(kpss_test_diff$p.value,2),
    Diff1_ADF_stat = round(adf_test_diff$statistic,2),
    Diff1_ADF_p_value = round(adf_test_diff$p.value,2)
  ))
}

# Printing results 
rownames(stationarity_results) <- NULL
print(stationarity_results)
```


Testing stationarity of differenced lifetime AMU trend time series 

```{r, warning=F, message=F}
# Defining AM-classes
am_group_names <- c(
  "Aminoglycosides",
  "Amphenicols",
  "SimPenicillins",
  "ExtPenicillins",
  "Lincosamides",
  "Macrolides",
  "Pleuromutilins",
  "Tetracyclines"
)

# Data frame to store stationarity test results 
stationarity_results <- data.frame(
  AMGroup = character(),
  Original_KPSS_stat = numeric(),
  Original_KPSS_p_value = numeric(),
  Original_ADF_stat = numeric(),
  Original_ADF_p_value = numeric(),
  Diff1_KPSS_stat = numeric(),
  Diff1_KPSS_p_value = numeric(),
  Diff1_ADF_stat = numeric(),
  Diff1_ADF_p_value = numeric(),
  stringsAsFactors = FALSE
)

# Loop to test stationarity individually for each AM-class for the lifetime AMU time series trend before and after differencing 
for (am in am_group_names) {

  # Filtering AM-class of lifetime AMU trend time series 
  data <- lifetime_AMU_trend %>%
    filter(am_group == am) %>%
    select(-am_group) 
  
  # Filtering AM-class of differenced lifetime AMU trend time series 
  data_diff <- lifetime_AMU_trend_diff %>%
    filter(am_group == am) %>%
    select(-am_group) 
  
  # Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test for trend stationarity
  kpss_test <- kpss.test(data$amu)
  kpss_test_diff <- kpss.test(data_diff$amu)
  
  # Augmented-Dickey–Fuller (ADF) test for trend stationarity
  adf_test <- adf.test(data$amu)
  adf_test_diff <- adf.test(data_diff$amu)
  
  # Storing stationarity results in data frame 
  stationarity_results <- rbind(stationarity_results, data.frame(
    AMGroup = am,
    Original_KPSS_stat = round(kpss_test$statistic,2),
    Original_KPSS_p_value = round(kpss_test$p.value,2),
    Original_ADF_stat = round(adf_test$statistic,2),
    Original_ADF_p_value = round(adf_test$p.value,2),
    Diff1_KPSS_stat = round(kpss_test_diff$statistic,2),
    Diff1_KPSS_p_value = round(kpss_test_diff$p.value,2),
    Diff1_ADF_stat = round(adf_test_diff$statistic,2),
    Diff1_ADF_p_value = round(adf_test_diff$p.value,2)
  ))
}

# Printing results 
rownames(stationarity_results) <- NULL
print(stationarity_results)
```

________________________________________________________________________________


ARG abundance summary statistics

```{r}
# Summarizing ARG abundance across all AMR-classes each month
ARG_abundance_trend_total <- ARG_abundance_trend %>%
  group_by(Date) %>%
  summarise(amr = sum(amr), .groups = "drop")

# Summary statistics for total ARG abundance each month
summary(ARG_abundance_trend_total)
```

```{r}
# Summarizing differenced (change in) ARG abundance across all AMR-classes each month
ARG_abundance_trend_diff_total <- ARG_abundance_trend_diff %>%
  group_by(Date) %>%
  summarise(amr = sum(amr), .groups = "drop")

# Summary statistics for differenced (change in) total ARG abundance each month
summary(ARG_abundance_trend_diff_total)
```


Lifetime AMU summary statistics

```{r}
# Summarizing lifetime AMU across all AM-classes each month
lifetime_AMU_trend_total <- lifetime_AMU_trend %>%
  group_by(Date) %>% 
  summarise(amu = sum(amu), .groups = "drop")

# Summary statistics for total lifetime AMU each month
summary(lifetime_AMU_trend_total)
```

```{r}
# Summarizing differenced (change in) lifetime AMU across all AM-classes each month
lifetime_AMU_trend_diff_total <- lifetime_AMU_trend_diff %>%
  group_by(Date) %>%
  summarise(amu = sum(amu), .groups = "drop")

# Summary statistics for differenced (change in) total lifetime AMU each month
summary(lifetime_AMU_trend_diff_total)
```
















































