
Title: Regression analysis: Dynamic regression time series models of ARG abundance with lifetime AMU as lagged predictor 
Author: Josefine Ostenfeld Nielsen, National Food Institute, Technical University of Denmark (josefineon@gmail.com)
Date: January 2024
Description: Computing 152 DR models and identifying models with positive and significant regression coefficients

________________________________________________________________________________


Installing packages and loading libraries

```{r, message=F, warning=F}
# Sourcing packages and libraries from the setup.R script
source(file.path("setup.R"))
```


Importing ARG abundance and lifetime AMU data

```{r}
# Reading ARG abundance and lifetime AMU files
ARG_abundance_trend_diff <- read.table(file.path("data/ARG_abundance_trend_diff.txt"), header = TRUE, sep = "\t")
lifetime_AMU_trend_diff <- read.table(file.path("data/lifetime_AMU_trend_diff.txt"), header = TRUE, sep = "\t")

# Formatting Date column 
ARG_abundance_trend_diff$Date <- as.Date(ARG_abundance_trend_diff$Date)
lifetime_AMU_trend_diff$Date <- as.Date(lifetime_AMU_trend_diff$Date)
```

________________________________________________________________________________


Computing DR models individually for each AM and corresponding AMR
OBS: approx. run time = 10 min (loop with automatic model selection of SARIMA errors for all 152 DR model)

```{r, message=F, warning = F}
# Defining AMR-class and AM-class in each model
plot_list <- list(
  list(column_amr = "Aminoglycoside", column_amu = "Aminoglycosides"),
  list(column_amr = "Amphenicol", column_amu = "Amphenicols"),
  list(column_amr = "Beta.lactam", column_amu = "SimPenicillins"),
  list(column_amr = "Beta.lactam", column_amu = "ExtPenicillins"),
  list(column_amr = "Lincosamide", column_amu = "Lincosamides"),
  list(column_amr = "Macrolide", column_amu = "Macrolides"),
  list(column_amr = "Pleuromutilin", column_amu = "Pleuromutilins"),
  list(column_amr = "Tetracycline", column_amu = "Tetracyclines"))

# Data frame to store DR models in each loop
DR_models <- data.frame(
  am_class = character(),
  lag = numeric(),
  reg_coef = numeric(),
  std = numeric(),
  p_val <- numeric(),
  model = character(),
  R = numeric(),
  AIC = numeric(),
  BIC = numeric(),
  r = numeric(),
  stringsAsFactors = FALSE
)

# Initialize an empty data frame all DR models
DR_models_all <- data.frame()

# Loop to compute DR model for each class, with nested loop to compute model with 0-18 months lagged lifetime AMU
for (n in plot_list) {
  
  # Filtering AM-class
  amu <- lifetime_AMU_trend_diff %>%
    filter(am_group == n$column_amu) %>%
    select(-am_group)
  
  # Filtering AMR-class
  amr <- ARG_abundance_trend_diff %>%
    filter(amr_class == n$column_amr) %>%
    select(-amr_class) 
  
  # Number of lags [months]
  lags <- seq(0, 18)
  
  # Loop to compute DR model for each class with 0-18 months lagged lifetime AMU
  for (n_lag in lags) {
    
    # Aligning AMR and time-lagged AMU time series for training data set (2015-2021)
    aligned_data <- amu %>%
      left_join(amr, by = "Date") %>%
      mutate(amu_lag = lag(amu, n = n_lag, default = NA)) %>%
      filter(!is.na(amr)) %>%
      filter(year(Date) < 2022) %>%
      mutate(Date = yearmonth(Date)) %>%
      as_tsibble(index = Date)
    
    # Computing DR model with automatic model selection to determine the order of p, q, P, and Q
    # Both AIC and BIC automatic model selection aplied, by running code twice, once with ic = "aic" and once with ic = "bic"
    model <- aligned_data %>%
      model(ARIMA(amr ~ 0 + amu_lag + pdq(d=0) + PDQ(D=0), ic = "aic"))
    
    # Estimating Pearson's correlation coefficient (one-sided)
    pearson_corr <- cor.test(aligned_data$amu_lag, aligned_data$amr, alternative = "greater", method = "pearson")
    
    # Estimating TSS (total sum of squares)
    tss <- sum((aligned_data$amr - mean(aligned_data$amr))^2)
    
    # Estimating RSS (residual sum of squares)
    fitted_values <- fitted(model)
    rss <- sum((aligned_data$amr - fitted_values$.fitted)^2)
    
    # Storing model output
    a <- tidy(model)
    
    # Storing DR model, coefficients, p-val, SARIMA terms order, R squared, model criterion, and correlation coefficient 
    DR_models <- rbind(DR_models, data.frame(
      am_class = n$column_amu,
      lag = n_lag,
      reg_coef = as.numeric(a[a$term == 'amu_lag', 'estimate']),
      std = as.numeric(a[a$term == 'amu_lag', 'std.error']),
      p_val = as.numeric(a[a$term == 'amu_lag', 'p.value']), 
      model = str_extract(model[[1]][1], "(?<=<).*(?=>)"),
      R = 1-(rss/tss),
      AIC = model[[1]][[1]][["fit"]][["fit"]][["AIC"]],
      BIC = model[[1]][[1]][["fit"]][["fit"]][["BIC"]],
      r = pearson_corr$estimate[["cor"]]
    ))
  }
  
  # Storing all DR models in combined data frame 
    if(nrow(DR_models_all) == 0){
      DR_models_all <- DR_models
    } else {
      DR_models_all <- rbind(DR_models_all, DR_models)
    }
  
  # Resetting data frame for storing DR models in each loop
  DR_models <- NULL
}

# Calculating one-sided p-value, filtering significant DR models
sign_DR_models <- DR_models_all %>%
  mutate(p_val = p_val/2)
  filter(p_val < 0.05) %>% 
  filter(reg_coef > 0) %>%
  filter(r > 0)

# Function to assign significance code 
assign_sig_code <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else if (p_value < 0.1) {
    return("˙")
  } else {
    return("")
  }
}

# Adding column with significant code
sign_DR_models$sign_code <- sapply(sign_DR_models$p_val, assign_sig_code)

# Printing models
print(sign_DR_models)
```


Significant DR models (all coefficients and diagnostic plots of innovation errors)

```{r}
# Identified significant DR models (un-commenting individually to run code separately for each model)

column_amr <- "Aminoglycoside"
column_amu <- "Aminoglycosides"
p1 <- 0
q1 <- 2
P1 <- 0
Q1 <- 1
lag <- 8

# column_amr <- "Amphenicol"
# column_amu <- "Amphenicols"
# p1 <- 0
# q1 <- 1
# P1 <- 0
# Q1 <- 0
# lag <- 8

# column_amr <- "Amphenicol"
# column_amu <- "Amphenicols"
# p1 <- 0
# q1 <- 1
# P1 <- 0
# Q1 <- 0
# lag <- 16

# column_amr <- "Beta.lactam"
# column_amu <- "ExtPenicillins"
# p1 <- 2
# q1 <- 1
# P1 <- 0
# Q1 <- 1
# lag <- 7

# column_amr <- "Beta.lactam"
# column_amu <- "ExtPenicillins"
# p1 <- 2
# q1 <- 1
# P1 <- 0
# Q1 <- 1
# lag <- 17

# column_amr <- "Beta.lactam"
# column_amu <- "SimPenicillins"
# p1 <- 2
# q1 <- 3
# P1 <- 0
# Q1 <- 0
# lag <- 7

# column_amr <- "Lincosamide"
# column_amu <- "Lincosamides"
# p1 <- 2
# q1 <- 0
# P1 <- 0
# Q1 <- 0
# lag <- 5

# column_amr <- "Macrolide"
# column_amu <- "Macrolides"
# p1 <- 0
# q1 <- 1
# P1 <- 0
# Q1 <- 0
# lag <- 0

# column_amr <- "Macrolide"
# column_amu <- "Macrolides"
# p1 <- 0
# q1 <- 1
# P1 <- 0
# Q1 <- 0
# lag <- 2

# column_amr <- "Pleuromutilin"
# column_amu <- "Pleuromutilins"
# p1 <- 3
# q1 <- 0
# P1 <- 0
# Q1 <- 1
# lag <- 7

# column_amr <- "Tetracycline"
# column_amu <- "Tetracyclines"
# p1 <- 0
# q1 <- 3
# P1 <- 0
# Q1 <- 1
# lag <- 9

# column_amr <- "Tetracycline"
# column_amu <- "Tetracyclines"
# p1 <- 0
# q1 <- 3
# P1 <- 1
# Q1 <- 2
# lag <- 14

# column_amr <- "Tetracycline"
# column_amu <- "Tetracyclines"
# p1 <- 0
# q1 <- 3
# P1 <- 0
# Q1 <- 1
# lag <- 15


# Filtering AM-class
amu <- lifetime_AMU_trend_diff %>%
  filter(am_group == column_amu) %>%
  select(-am_group)

# Filtering AMR-class
amr <- ARG_abundance_trend_diff %>%
  filter(amr_class == column_amr) %>%
  select(-amr_class) 

# Aligning AMR and time-lagged AMU (identified significant predictor) time series for training data set (2015-2021)
aligned_data <- amu %>%
  left_join(amr, by = "Date") %>%
  mutate(amu_lag = lag(amu, n = lag, default = NA)) %>%
  filter(!is.na(amr)) %>%
  filter(year(Date) < 2022) %>%
  mutate(Date = yearmonth(Date)) %>%
  as_tsibble(index = Date)

# Computing identified significant DR model
model <- aligned_data %>%
  model(ARIMA(amr ~ 0 + amu_lag + pdq(p1,0,q1) + PDQ(P1,0,Q1)))

# Printing model output
report(model)
tidy(model)

# Regression errors diagnostic plots
residuals(model, type = "regression") %>%
  gg_tsdisplay(.resid, plot_type = "scatter") +
  labs(title = "Regression errors")

# Innovation errors diagnostic plots
residuals(model, type = "innovation") %>%
  gg_tsdisplay(.resid, plot_type = "scatter") +
  labs(title = "Innovation errors")
```


