
Title: Manuscript figures
Author: Josefine Ostenfeld Nielsen
Date: January 2024
Description: Figures for the manuscript and supplementary material 

________________________________________________________________________________


Installing packages and loading libraries

```{r, message=F, warning=F}
# Sourcing packages and libraries from the setup.R script 
source(file.path("setup.R"))
```


Importing ARG abundance time series data

```{r}
# Reading data
ARG_abundance <- read.table(file.path("data/ARG_abundance.txt"), header = TRUE, sep = "\t")
ARG_abundance_interpolated <- read.table(file.path("data/ARG_abundance_int.txt"), header = TRUE, sep = "\t")
ARG_abundance_adjusted <- read.table(file.path("data/ARG_abundance_adj.txt"), header = TRUE, sep = "\t")
ARG_abundance_trend <- read.table(file.path("data/ARG_abundance_trend.txt"), header = TRUE, sep = "\t")
ARG_abundance_trend_diff <- read.table(file.path("data/ARG_abundance_trend_diff.txt"), header = TRUE, sep = "\t")

lifetime_AMU <- read.table(file.path("data/lifetime_AMU.txt"), header = TRUE, sep = "\t")
lifetime_AMU_trend <- read.table(file.path("data/lifetime_AMU_trend.txt"), header = TRUE, sep = "\t")
lifetime_AMU_trend_diff <- read.table(file.path("data/lifetime_AMU_trend_diff.txt"), header = TRUE, sep = "\t")


# Formatting Date column 
ARG_abundance$Date <- as.Date(ARG_abundance$Date)
ARG_abundance_adjusted$Date <- as.Date(ARG_abundance_adjusted$Date)
ARG_abundance_interpolated$Date <- as.Date(ARG_abundance_interpolated$Date)
ARG_abundance_trend$Date <- as.Date(ARG_abundance_trend$Date)
ARG_abundance_trend_diff$Date <- as.Date(ARG_abundance_trend_diff$Date)

lifetime_AMU$Date <- as.Date(lifetime_AMU$Date)
lifetime_AMU_trend$Date <- as.Date(lifetime_AMU_trend$Date)
lifetime_AMU_trend_diff$Date <- as.Date(lifetime_AMU_trend_diff$Date)
```

________________________________________________________________________________


ARG abundance time series data (after interpolation, before adjustment of systematic error and decomposition)

```{r, warning=F, fig.height=9, fig.width=8}
# Plotting ARG abundance time series
ggplot() +
  geom_line(data = ARG_abundance_interpolated, aes(y = amr, x = Date), color = "#1A237E", size = 1) +
  facet_grid(rows = vars(amr_class), scales = "free") +
  coord_cartesian(xlim = as.Date(c("2015-01-01", "2023-01-01"))) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
  labs(title = "", x = "", y = "ARG abundance [FPKM]", color = "Source") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.spacing = unit(0.75, "lines")
  ) +
  geom_hline(yintercept = c(0, NA), color = "lightgray") +
  geom_vline(xintercept = as.Date(seq(from = as.Date("2015-01-01"), to = as.Date("2023-01-01"), by = "years")), color = "lightgray")
```


Unadjusted and adjusted ARG abundance time series trends

```{r, message=FALSE,warning = FALSE, fig.width=8, fig.height=4}
# Defining AMR-classes
amr_class_names <- c(
  "Aminoglycoside",
  "Amphenicol",
  "Beta.lactam",
  "Lincosamide",
  "Macrolide",
  "Pleuromutilin",
  "Tetracycline"
)

# Function for z-score standardization
standardize_columns <- function(data) {
  numeric_columns <- sapply(data, is.numeric)
  data[, numeric_columns] <- lapply(data[, numeric_columns, drop = FALSE], function(x) {
    as.numeric(scale(x))
  })
  return(data)
}

# Data frame to store z-score normalized ARG abundance time series 
ARG_abundance_interpolated_z <- data.frame(
  Date = as.Date(character()),
  amr_class = character(),
  amr = numeric(),
  stringsAsFactors = FALSE
)

# Loop to isolate trend component and z-score normalize ARG abundance for each AMR-class
for (class in amr_class_names) {
  
  # Filtering AMR-class
  amr <- ARG_abundance_interpolated %>%
    filter(amr_class == class) %>%
    select(-amr_class)
  
  # Converting to tsibble object (time series Date format)
  amr_ts <- amr %>%
    mutate(Date = yearmonth(Date)) %>%
    as_tsibble(index = Date)
  
  # Decomposing and isolating trend component
  amr_decomposed <- stl(amr_ts, s.window = "periodic", t.window = 7)
  amr_decomposed <- data.frame(amr_decomposed$time.series)
  amr_trend <- bind_cols(Date = amr$Date, amr = amr_decomposed$trend)
  
  # Removing first and last two values where smoothing was incomplete, and z-score standardizing 
  amr_trend <- amr_trend %>%
    standardize_columns() %>%
    mutate(
      amr_class = class
    ) %>%
    slice(3:(n()-2))
  
  # Saving z-score standardized ARG abundance trends 
  ARG_abundance_interpolated_z <- rbind(ARG_abundance_interpolated_z, amr_trend)
}  
  
# Data frame to store z-score normalized adjusted ARG abundance time series   
ARG_abundance_adjusted_z <- data.frame(
  Date = as.Date(character()),
  amr_class = character(),
  amr = numeric(),
  stringsAsFactors = FALSE
)
  
# Loop to isolate trend component and z-score normalize adjusted ARG abundance for each AMR-class
for (class in amr_class_names) {
  
  # Filtering AMR-class
  amr <- ARG_abundance_adjusted %>%
    filter(amr_class == class) %>%
    select(-amr_class) 
  
  # Converting to tsibble object (time series Date format)
  amr_ts <- amr %>%
    mutate(Date = yearmonth(Date)) %>%
    as_tsibble(index = Date)
  
  # Decomposing and isolating trend component
  amr_decomposed <- stl(amr_ts, s.window = "periodic", t.window = 7)
  amr_decomposed <- data.frame(amr_decomposed$time.series)
  amr_trend <- bind_cols(Date = amr$Date, amr = amr_decomposed$trend)
  
  # Removing first and last two values where smoothing was incomplete, and z-score standardizing 
  amr_trend <- amr_trend %>%
    standardize_columns() %>%
    mutate(
      amr_class = class
    ) %>%
    slice(3:(n()-2))
  
  # Saving z-score standardized adjusted ARG abundance trends 
  ARG_abundance_adjusted_z <- rbind(ARG_abundance_adjusted_z, amr_trend)
}  
  
# Defining AMR-class names for plot legend
custom_labels <- c("Aminoglycoside" = "Aminoglycoside", 
                   "Amphenicol" = "Amphenicol", 
                   "Beta.lactam" = "Beta lactam", 
                   "Lincosamide" = "Lincosamide",
                   "Macrolide" = "Macrolide", 
                   "Pleuromutilin" = "Pleuromutilin", 
                   "Tetracycline" = "Tetracycline")

# Plot of z-score normalized ARG abundance trends (before adjustment)
ggplot() +
    geom_line(data = ARG_abundance_interpolated_z, aes(x = Date, y = amr, color = amr_class), size = 1) +
    coord_cartesian(xlim = as.Date(c("2015-01-01", "2023-01-01")), ylim = c(-3,3)) +
    scale_y_continuous(name = "Z-score", expand = c(0, 0)) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
    labs(title = "ARG abundance trends (unadjusted)", x = "") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
      axis.text.y = element_text(size = 10),
      panel.spacing = unit(0.75, "lines"),
      panel.border = element_rect(color = "lightgray", fill = NA, size = 1),
      panel.grid.minor = element_blank(),
      title = element_text(size=12),
      legend.position = "bottom",
      legend.text = element_text(size = 12)
    ) +
  scale_color_discrete(name = "AMR class", labels = custom_labels)

# Plot of z-score normalized ARG abundance trends (after adjustment)
ggplot() +
    geom_line(data = ARG_abundance_adjusted_z, aes(x = Date, y = amr, color = amr_class), size = 1) +
    coord_cartesian(xlim = as.Date(c("2015-01-01", "2023-01-01")), ylim = c(-3,3)) +
    scale_y_continuous(name = "Z-score", expand = c(0, 0)) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
    labs(title = "ARG abundance trends (adjusted)", x = "") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
      axis.text.y = element_text(size = 10),
      panel.spacing = unit(0.75, "lines"),
      panel.border = element_rect(color = "lightgray", fill = NA, size = 1),
      panel.grid.minor = element_blank(),
      title = element_text(size=12),
      legend.position = "bottom",
      legend.text = element_text(size = 12)
    ) +
  scale_color_discrete(name = "AMR class", labels = custom_labels)
```


Correlation matrix plot of the correlation between the ARG abundances between each AMR-class

```{r, fig.width=10, fig.height=10, warning=FALSE}
# Preparing ARG abundance data frame
corr_data <- ARG_abundance_adjusted %>%
  pivot_wider(names_from = amr_class, values_from = amr) %>%
  select(-Date)

# Custom panel function for correlation coefficients
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor = 1.25, col.cor = "#1A237E") {
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y, use = "complete.obs")
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  text(0.5, 0.5, txt, cex = cex.cor, col = col.cor)
}

# Custom panel function for scatter plots
panel.scatter <- function(x, y, col = "#1A237E") {
  points(x, y, pch = 19, col = col, cex = 0.8)
}

# Correlation matrix plot 
pairs(corr_data, 
      upper.panel = panel.cor,
      lower.panel = panel.scatter,
      cex.labels = 1.25)
```


ARG abundance time series decomposition plots

```{r, warning=FALSE, fig.width=7, fig.height=6}
# Defining AMR-class names
am_group_names <- c(
  "Aminoglycoside",
  "Amphenicol",
  "Beta.lactam",
  "Lincosamide",
  "Macrolide",
  "Pleuromutilin",
  "Tetracycline"
)

# Loop ti decompose and plot time series decomposition plot of the ARG abundance for each AMR-class 
for (am in am_group_names) {
  
  # Filtering AMR-class and converting into tsibble object (time series Date format)
  data <- ARG_abundance_adjusted %>%
    filter(amr_class == am) %>%
    select(-amr_class) %>%
    mutate(Date = yearmonth(Date)) %>%
    as_tsibble(index = Date)
  
  # Time series decomposition
  decomposed <- stl(data, s.window = "periodic", t.window = 7)
  
  # Decomposition plot
  plot(decomposed, main = paste("ARG abundance: Classical additive decomposition of", am))
}
```


Time series trend plots of corresponding AMU and AMR

```{r, message=FALSE, fig.height=12, fig.width=10, warning = FALSE}
# Defining AM-class and corresponding AMR-class
plot_list <- list(
  list(column_amr = "Aminoglycoside", column_amu = "Aminoglycosides", title = "Aminoglycoside"),
  list(column_amr = "Amphenicol", column_amu = "Amphenicols", title = "Amphenicol"),
  list(column_amr = "Beta.lactam", column_amu = "ExtPenicillins", title = "Beta lactam/Extended-spectrum penicillin"),
  list(column_amr = "Beta.lactam", column_amu = "SimPenicillins", title = "Beta lactam/Narrow-spectrum penicillin"),
  list(column_amr = "Lincosamide", column_amu = "Lincosamides", title = "Lincosamide"),
  list(column_amr = "Macrolide", column_amu = "Macrolides", title = "Macrolide"),
  list(column_amr = "Pleuromutilin", column_amu = "Pleuromutilins", title = "Pleuromutilin"),
  list(column_amr = "Tetracycline", column_amu = "Tetracyclines", title = "Tetracycline"))

# List to store each plot of the corresponding AMU and AMR
plot_grid_ts <- list()

# Loop to compute each corresponding AMU and AMR time series trend plot and store in list 
for (n in plot_list) {
  
  # Initiating list and storing as the next object in list in each loop
  plot_grid_ts[[length(plot_grid_ts) + 1]] <- local({
  
    # Filtering AM-class
    amu <- lifetime_AMU_trend %>%
      filter(am_group == n$column_amu) %>%
      select(-am_group)
    
    # Filtering AMR-class
    amr <- ARG_abundance_trend %>%
      filter(amr_class == n$column_amr) %>%
      select(-amr_class) 
    
    # Re-scale factor to rescale ARG abundance values for secondary y-axis
    rescale_factor <- mean(amr$amr)/mean(amu$amu)
    
    # Re-scaling ARG abundance values for secondary y-axis
    amr$amr_scaled <- amr$amr/rescale_factor

    # Defining primary y-axis limits
    max_val <- max(max(amu$amu, na.rm = TRUE), max(amr$amr_scaled, na.rm = TRUE))
    min_val <- min(min(amu$amu, na.rm = TRUE), min(amr$amr_scaled, na.rm = TRUE))
    
    # Defining legend text
    color1 <- "Lifetime AMU"
    color2 <- "ARG abundance"
    
    # Plotting time series trends
    p_ts <- ggplot() +
      geom_line(data = amu, aes(x = Date, y = amu, color = color1), size = 0.75, alpha = 0.75) +
      geom_line(data = amr, aes(x = Date, y = amr_scaled, color = color2), size = 0.75) +
      scale_color_manual(values = c("blue", "red"), breaks = c(color1, color2)) +
      coord_cartesian(xlim = as.Date(c("2013-01-01", "2023-01-01"))) +
      scale_y_continuous(
        name = "Lifetime AMU [#ADDkg/pig]",
        sec.axis = sec_axis(trans = ~ . , labels = scales::number_format(scale = rescale_factor),name = "ARG abundance [FPKM]")
      ) +
      scale_x_date(
        name = "",
        date_breaks = "1 year", 
        minor_breaks = "1 year",
        date_labels = "%Y", 
        expand = c(0, 0),
      ) +
      labs(title = n$title, color = "") +
      theme_minimal() +
      theme(
        axis.text.x.bottom = element_text(angle = 90, hjust = 1, vjust = 0.2, size = 12),
        axis.text.y.left = element_text(size = 12),
        axis.text.y.right = element_text(size = 12),
        axis.title.x.bottom = element_text(size = 12),
        axis.title.x.top = element_text(size = 12),
        panel.spacing = unit(0.75, "lines"),
        panel.border = element_rect(color = "lightgray", fill = NA, size = 1),
        title = element_text(size=12),
        plot.title = element_text(size=12),
        legend.text = element_text(size = 12),
        legend.position = "top")
    
    # Storing plot in list
    return(p_ts)
  })
}

# Extracting the legend from the first plot 
legend <- get_legend(plot_grid_ts[[1]])

# Removing axis titles and legend from all plots stored in list 
plots <- lapply(plot_grid_ts, function(p) {
  p + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none")})

# Making combined plot with each corresponding class time series trends plots, adding common axis titles and legend
plots_axis <- grid.arrange(
  grobs = plots,
  ncol = 2,
  nrow = 4,
  widths = c(1, 1),
  bottom = legend,
  left = textGrob("Lifetime AMU [#ADDkg/pigs]", rot = 90, gp = gpar(fontsize = 12)),
  right = textGrob("ARG abundance [FPKM]", rot = -90, gp = gpar(fontsize = 12))
)
```


Differenced time series trend plots of corresponding AMU and AMR

```{r, message=FALSE, fig.height=11, fig.width=10, warning = FALSE}
# Defining AM-class and corresponding AMR-class
plot_list <- list(
  list(column_amr = "Aminoglycoside", column_amu = "Aminoglycosides", title = "Aminoglycoside"),
  list(column_amr = "Amphenicol", column_amu = "Amphenicols", title = "Amphenicol"),
  list(column_amr = "Beta.lactam", column_amu = "ExtPenicillins", title = "Beta lactam/Extended-spectrum penicillin"),
  list(column_amr = "Beta.lactam", column_amu = "SimPenicillins", title = "Beta lactam/Narrow-spectrum penicillin"),
  list(column_amr = "Lincosamide", column_amu = "Lincosamides", title = "Lincosamide"),
  list(column_amr = "Macrolide", column_amu = "Macrolides", title = "Macrolide"),
  list(column_amr = "Pleuromutilin", column_amu = "Pleuromutilins", title = "Pleuromutilin"),
  list(column_amr = "Tetracycline", column_amu = "Tetracyclines", title = "Tetracycline"))

# Function for z-score standardization
standardize_columns <- function(data) {
  numeric_columns <- sapply(data, is.numeric)
  data[, numeric_columns] <- lapply(data[, numeric_columns, drop = FALSE], function(x) {
    as.numeric(scale(x))
  })
  return(data)
}

# List to store each plot of the corresponding AMU and AMR
plot_grid_ts <- list()

# Loop to compute each corresponding AMU and AMR differenced time series trend plot and store in list 
for (n in plot_list) {
  
  # Initiating list and storing as the next object in list in each loop
  plot_grid_ts[[length(plot_grid_ts) + 1]] <- local({
  
    # Filtering AM-class and z-score standardizing the differenced lifetime AMU
    amu <- lifetime_AMU_trend_diff %>%
      filter(am_group == n$column_amu) %>%
      select(-am_group) %>%
      standardize_columns()
    
    # Filtering AMR-class and z-score standardizing the differenced ARG abundance
    amr <- ARG_abundance_trend_diff %>%
      filter(amr_class == n$column_amr) %>%
      select(-amr_class) %>%
      standardize_columns()
    
    # Defining legend text
    color1 <- "Lifetime AMU´"
    color2 <- "ARG abundance´"
    
    # Plotting time series trends
    p_ts <- ggplot() +
      geom_line(data = amu, aes(x = Date, y = amu, color = color1), size = 0.75, alpha = 0.75) +
      geom_line(data = amr, aes(x = Date, y = amr, color = color2), size = 0.75) +
      scale_color_manual(values = c("blue", "red"), breaks = c(color1, color2)) +
      coord_cartesian(xlim = as.Date(c("2013-01-01", "2023-01-01"))) +
      scale_y_continuous(name = "Z-score", limits = c(-5,5), expand = c(0,0)) +
      scale_x_date(name = "", date_breaks = "1 year", minor_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
      labs(title = n$title, color = "") +
      theme_minimal() +
      theme(
        axis.text.x.bottom = element_text(angle = 90, hjust = 1, vjust = 0.2, size = 12),
        axis.text.y.left = element_text(size = 12),
        axis.text.y.right = element_text(size = 12),
        axis.title.x.bottom = element_text(size = 12),
        axis.title.x.top = element_text(size = 12),
        panel.spacing = unit(0.75, "lines"),
        panel.border = element_rect(color = "lightgray", fill = NA, size = 1),
        title = element_text(size=12),
        plot.title = element_text(size=12),
        legend.text = element_text(size = 12),
        legend.position = "top")
    
    # Storing plot in list
    return(p_ts)
  })
}

# Extracting the legend from the first plot 
legend <- get_legend(plot_grid_ts[[1]])

# Removing axis titles and legend from all plots stored in list 
plots <- lapply(plot_grid_ts, function(p) {
  p + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none")})

# Making combined plot with each corresponding class time series trends plots, adding common axis titles and legend
plots_axis <- grid.arrange(
  grobs = plots,
  ncol = 2,
  nrow = 4,
  bottom = legend,
  left = textGrob("Z-score", rot = 90, gp = gpar(fontsize = 12))
)
```


Pearson's correlation coefficient (r) heatmap for each corresponding AMR and AMU at lag 0 - 18 months

```{r, message=FALSE,warning = FALSE, fig.height=4, fig.width=9}
# Defining max time lag of lifetime AMU
lag_max <- 18

# Defining AM-class and corresponding AMR-class
plot_list <- list(
  list(column_amr = "Aminoglycoside", column_amu = "Aminoglycosides", title = "Aminoglycoside"),
  list(column_amr = "Amphenicol", column_amu = "Amphenicols", title = "Amphenicol"),
  list(column_amr = "Beta.lactam", column_amu = "ExtPenicillins", title = "B.lac/E.Pen."),
  list(column_amr = "Beta.lactam", column_amu = "SimPenicillins", title = "B.lac/N.Pen."),
  list(column_amr = "Lincosamide", column_amu = "Lincosamides", title = "Lincosamide"),
  list(column_amr = "Macrolide", column_amu = "Macrolides", title = "Macrolide"),
  list(column_amr = "Pleuromutilin", column_amu = "Pleuromutilins", title = "Pleuromutilin"),
  list(column_amr = "Tetracycline", column_amu = "Tetracyclines", title = "Tetracycline"))

# Function to assign sign. code
assign_sig_code <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else if (p_value < 0.1) {
    return("˙")
  } else {
    return("")
  }
}

# Data frame to store results from correlation test
ccf_df <- data.frame(
  amr_class = character(),
  lag = numeric(),
  corr = numeric(),
  p_val = numeric(),
  stringsAsFactors = FALSE
)

# Initialize an empty data frame for combined results for all classes for the heatmap
hm_df <- data.frame()

# Loop to estimate Pearson's correlation coefficient for each corresponding AMR and AMU at each lag
for (n in plot_list) {
  
  # Filtering AM-class
  amu <- lifetime_AMU_trend_diff %>%
    filter(am_group == n$column_amu) %>%
    select(-am_group)
  
  # Filtering AMR-class
  amr <- ARG_abundance_trend_diff %>%
    filter(amr_class == n$column_amr) %>%
    select(-amr_class) 
  
  # Number of lags to be investigated
  lags <- seq(0, lag_max)
  
  # Nested loop to make cor.test for each lag individually 
  for (n_lag in lags) {
    
    # Aligning AMR and time-lagged AMU time series for training data set (2015-2021)
    aligned_data <- amu %>%
      left_join(amr, by = "Date") %>%
      mutate(amu = lag(amu, n = n_lag, default = NA)) %>%
      filter(!is.na(amr)) %>%
      filter(year(Date) < 2022) 
    
    # Estimating correlation coefficient
    ccf_results <- cor.test(aligned_data$amu, aligned_data$amr, alternative = "greater", method = "pearson")
  
    # Storing results 
    ccf_df <- rbind(ccf_df, data.frame(
      amr_class = n$title,
      lag = n_lag,
      corr = ccf_results$estimate[["cor"]],
      p_val = ccf_results$p.value
    ))
  }
  
  # Assigning significance code to p-value
  ccf_df$sign_code <- sapply(ccf_df$p_val, assign_sig_code)
  
  # Storing results in compbined results data frame for the heatmap
  if(nrow(hm_df) == 0){
    hm_df <- ccf_df
  } else {
    hm_df <- rbind(hm_df, ccf_df)
  }
  
  # Resetting data frame for storing correlation results in each loop
  ccf_df <- NULL
}

# Alphabetically sorting results for each class for visualization purposes in the heatmap
hm_df$amr_class <- factor(hm_df$amr_class, levels = rev(sort(unique(hm_df$amr_class))))

# Computing heatmap of correlation coefficients
ggplot(hm_df, aes(x = lag, y = amr_class, fill = corr)) +
  geom_tile(color = "white") +  
  geom_text(aes(label = sign_code), color = "black", angle = 0, vjust = 0.75) +  
  scale_fill_gradient2(low = "blue", high = "red", mid = "#eeeeee", midpoint = 0, limit = c(-0.5, 0.5), space = "Lab") +  
  scale_x_continuous(limits = c(-0.5,lag_max+0.5), breaks = seq(0, lag_max, by = 1), expand = c(0,0)) +
  theme_minimal() +  
  theme(
    axis.text.x = element_text(angle = 0, hjust = 1),  
    plot.title = element_text(hjust = -0.25, size = 10)
  ) +
  labs(title = "", x = "Lag [months]", y = "", fill = "r") 


```


ARG abundance 2022 forecast plot

```{r, message=FALSE, fig.height=12, fig.width=10, warning = FALSE}
# Defining AMU/AMR combinations and titles
plot_list <- list(
  list(column_amr = "Aminoglycoside", column_amu = "Aminoglycosides", title = "Aminoglycoside (lag: 8 months)", p = 0, q = 2, Q = 1, lag = 8),
  list(column_amr = "Amphenicol", column_amu = "Amphenicols", title = "Amphenicol (lag: 8 months)", p = 0, q = 1, Q = 0, lag = 8),
  list(column_amr = "Beta.lactam", column_amu = "ExtPenicillins", title = "Beta lactam/Extended-spectrum penicillins (lag: 7 months)", p = 2, q = 1, Q = 1, lag = 7),
  list(column_amr = "Beta.lactam", column_amu = "SimPenicillins", title = "Beta lactam/Narrow-spectrum penicillins (lag: 7 months)", p = 2, q = 3, Q = 0, lag = 7),
  list(column_amr = "Lincosamide", column_amu = "Lincosamides", title = "Lincosamide (lag: 5 months)", p = 2, q = 0, Q = 0, lag = 5),
  list(column_amr = "Macrolide", column_amu = "Macrolides", title = "Macrolide (no lag)", p = 0, q = 1, Q = 0, lag = 0),
  list(column_amr = "Pleuromutilin", column_amu = "Pleuromutilins", title = "Pleuromutilin (lag: 7 months)", p = 3, q = 0, Q = 1, lag = 7),
  list(column_amr = "Tetracycline", column_amu = "Tetracyclines", title = "Tetracycline (lag: 9 months)", p = 0, q = 3, Q = 1, lag = 9))

# List to store plots
plot_grid_ts <- list()

# 
for (n in plot_list) {
  
  # Initiating list and storing as the next object in list in each loop
  plot_grid_ts[[length(plot_grid_ts) + 1]] <- local({
    
      # Filtering AM-class
      amu <- lifetime_AMU_trend_diff %>%
        filter(am_group == n$column_amu) %>%
        select(-am_group)
      
      # Filtering AMR-class
      amr <- ARG_abundance_trend_diff %>%
        filter(amr_class == n$column_amr) %>%
        select(-amr_class) 
      
      # Converting into tsibble object (time series Date format)
      amu_trend <- lifetime_AMU_trend %>%
        filter(am_group == n$column_amu) %>%
        select(-am_group)
      
      # Converting into tsibble object (time series Date format)
      amr_trend <- ARG_abundance_trend %>%
        filter(amr_class == n$column_amr) %>%
        select(-amr_class) 
    
      # Aligning AMR and time-lagged AMU time series for training data set (2015-2021)
      aligned_data <- amu %>%
        left_join(amr, by = "Date") %>%
        mutate(amu = lag(amu, n = n$lag, default = NA)) %>%
        filter(!is.na(amr)) %>%
        filter(year(Date) < 2022) %>%
        mutate(Date = yearmonth(Date)) %>%
        as_tsibble(index = Date)
      
      # Computing identified significant DR model
      model <- aligned_data %>%
        model(ARIMA(amr ~ 0 + amu + pdq(n$p,0,n$q) + PDQ(0,0,n$Q)))
      
      # Defining the subset of the lifetime AMU time series identified as the significant predictor of the ARG abundance from January 2022 to October 2022
      data_forecast <- amu %>%
        slice((n()-(11+n$lag)):(n()-(2+n$lag))) %>%
        mutate(Date = Date + months(n$lag)) %>%
        mutate(Date = yearmonth(Date)) %>%
        as_tsibble(index = Date)
      
      # Forecasting ARG abundance 2022 trend
      fcast <- forecast(model, new_data = data_forecast)
      
      # Storing forecast (chance in ARG abundance from the month prior) with respective months in 2022
      fcast_result <- bind_cols(Date = amu[(nrow(amu)-11):(nrow(amu)-2), 1], amr_diff_forecast = fcast$.mean)
      
      # Initiating vector to store absolute ARG abundance forecast values for the 10 months (January 2022 to October 2022) 
      fcast_result$amr_forecast <- numeric(10)
      
      # Estimating absolute ARG abundance forecasted for January 2022
      fcast_result$amr_forecast[1] <- amr_trend$amr[80] + fcast_result$amr_diff_forecast[1]
    
      # Loop through the remaining rows to calculate absolute ARG abundance forecast values for each month
      for(i in 2:nrow(fcast_result)) {
        fcast_result$amr_forecast[i] <- fcast_result$amr_forecast[i-1] + fcast_result$amr_diff_forecast[i]
      }
      
      # Defining ARG abundance training dataset
      training_data <- amr_trend %>% filter(year(Date) < 2022) 
      
      # Defining ARG abundance test dataset 
      test_data <- amr_trend %>% filter(year(Date) > 2019)
      
      # Re-scale factor to rescale ARG abundance values for secondary y-axis
      amr <- amr_trend %>% filter(year(Date) > 2019)
      amu <- amu_trend %>% filter(year(Date) > 2019)
      rescale_factor <- mean(amr$amr)/mean(amu$amu)
      
      # Re-scaling ARG abundance values for secondary y-axis
      amr$amr_scaled <- amr$amr/rescale_factor
      fcast_result$amr_forecast_scaled <- fcast_result$amr_forecast/rescale_factor
      
      # Defining primary y-axis limits
      max_val <- max(max(amu$amu, na.rm = TRUE), max(amr$amr_scaled, na.rm = TRUE), max(fcast_result$amr_forecast_scaled, na.rm = TRUE))
      min_val <- min(min(amu$amu, na.rm = TRUE), min(amr$amr_scaled, na.rm = TRUE), min(fcast_result$amr_forecast_scaled, na.rm = TRUE))
      
      # Defining legend text
      color1 <- "Lifetime AMU"
      color2a <- "ARG abundance (training data)"
      color2b <- "ARG abundance (test data)"
      color3 <- "ARG abundance (forecast)"
      
            # Computing forecast plot
      p_ts <- ggplot() +
        geom_line(data = amu_trend, aes(x = Date, y = amu, color = color1), size = 1, alpha = 0.5) +
        geom_line(data = test_data, aes(x = Date, y = amr/rescale_factor, color = color2b), linetype = "twodash", size = 1) +
        geom_line(data = training_data, aes(x = Date, y = amr/rescale_factor, color = color2a), size = 1) +
        geom_line(data = fcast_result, aes(x = Date, y = amr_forecast/rescale_factor, color = color3), linetype = "twodash", size = 1) +
        scale_color_manual(values = c("blue", "red", "red", "black"), breaks = c(color1, color2a, color2b, color3)) +
        guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "twodash", "twodash"))), linetype = FALSE) +
        labs(title = n$title, x = "", y = "", color = "") +
        scale_y_continuous(
          name = "Lifetime AMU [#ADDkg/pig]", limits = c(min_val, max_val),
          sec.axis = sec_axis(trans = ~ . , labels = scales::number_format(scale = rescale_factor), name = "ARG abundnce [FPKM]"),
        ) +
        scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
        coord_cartesian(xlim = as.Date(c("2020-01-01", "2023-01-01"))) +
        theme_minimal() +
        theme(
          axis.text.x.bottom = element_text(angle = 90, hjust = 1, vjust = 0.2, size = 12),
          axis.text.y.left = element_text(size = 12),
          axis.text.y.right = element_text(size = 12),
          axis.title.x.bottom = element_text(size = 12),
          axis.title.x.top = element_text(size = 12),
          panel.spacing = unit(0.75, "lines"),
          panel.border = element_rect(color = "lightgray", fill = NA, size = 1),
          title = element_text(size=12),
          plot.title = element_text(size=12),
          legend.text = element_text(size = 12),
          legend.position = "top")
      
      # Storing plot in list  
      return(p_ts)
    })
}

# Extracting the legend from the first plot 
legend <- get_legend(plot_grid_ts[[1]])

# Removing axis titles and legend from all plots stored in list 
plots <- lapply(plot_grid_ts, function(p) {
  p + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none")})

# Making combined plot with each corresponding class time series trends plots, adding common axis titles and legend
plots_axis <- grid.arrange(
  grobs = plots,
  ncol = 2,
  nrow = 4,
  bottom = legend,
  left = textGrob("Lifetime AMU [#ADDkg/pigs]", rot = 90, gp = gpar(fontsize = 12)),
  right = textGrob("ARG abundance [FPKM]", rot = -90, gp = gpar(fontsize = 12))
)
```





